---
title: "Building and deploying an ML model for patient readmissions predictions"
subtitle: "MLOps using R, tidymodels, docker, and Vetiver."
date: "`r Sys.Date()`"
output: html_document
author: "Dan A. Tshisungu"
---

# 1. About Data Analysis Report
This RMarkdown file contains the report of the data analysis done for the project on MLOps in R: Deploying machine learning models using Vetiver. It contains the tasks and steps to complete the project, such as building a stacked prediction model, storing and version the model, and deploying the model. The final report was completed on `r date()`. 

**Context:**

Did you know that hospital readmissions within 30 days cost the U.S. healthcare system over $17 billion on average annually? 

Having build a prediction model for hospital readmission, the next phase would be to deploy the model for clinical use.

**Project artifact:**

By the end of the project, learners will have constructed a fully functional deployment pipeline using Vetiver in R and monitoring tools to deploy and manage machine learning models effectively.

**Data Description:**

The data set for this project is an excerpt of the dataset provided during the Visual Automated Disease Analytics (VADA) summer school training, 2018. The VADA Summer School training dataset was derived from the Health Facts database (Cerner Corporation, Kansas City, MO, USA). This database contains clinical records from 130 participating hospitals across the USA. These clinical records contain information pertaining to 69,984 observations and 27 variables including patient encounter data, demographics, HbA1c levels, diagnostic testing and treatments, and patient outcomes. Data used were from 1999–2008 from a cohort of 130 hospitals, deidentified and trimmed to include only inpatient visits.

Strack B, DeShazo JP, Gennings C, et al. Impact of HbA1c measurement on hospital readmission rates: analysis of 70,000 clinical database patient records. Biomed Res Int. 2014;2014:781670. doi:10.1155/2014/781670


# 2. Setup and data loading

In this task, you will import the required packages and data for this project. You may need to install these packages if not installed. You can use the `install.packages()` function.

```{r data message = F, warning = F}
## Import required packages
library(tidyverse)
library(tidymodels)
library(themis)
library(vetiver)
library(plumber)
library(gridExtra)
library(skimr)
library(DataExplorer)
library(gt)
library(stacks)
library(pins)
library(here)

## Load the data set
here::i_am("Deploy_ML_models_using_vetiver.Rmd")

readmit_df <- read_csv(here("data", "final_readmit_df.csv"))

## Quick data pre-processing
readmit_df <- readmit_df %>% 
  ## Convert character variables into factors
  mutate_at(vars(!c(date, hospital_stay, patient_visits, 
                    num_medications, num_diagnosis)), as.factor)
 

## Take a glimpse at the data
glimpse(readmit_df)
```

# 3. Exploratory Data Analysis

## 3.1. Summary 

```{r description}
introduce(readmit_df) %>% 
  gt()
```

**Note:**
- There are no missing value.

## 3.2. Frequency distribution

```{r frequency}
plot_bar(readmit_df,
         title = "Frequency distribution of categorical variables")
```


**Note:**

- Our target variable, `readmitted` is not very imbalanced which should give us less work.

## 3.3. Numerical features distributions

```{r num_featu}
plot_histogram(readmit_df,
               title = "Distributions of numerical features")
```
**Note**
 
 -Numerical features are skewed, informing a need for transformation.
 
## 3.4. Correlation between numerical features
```{r correlation}

plot_correlation(readmit_df,
                 type = "continuous",
                 title = "Correlation between continuous features")
```
 
**Note:**

- Features are not correlated between them, which is good.


# 4. Data splitting

## 4.1. Create a validation set

```{r val set}
## Create a sample of 300 rows from the data
readmit_val_set <-
  readmit_df %>% 
  slice_tail(n = 300) |> 
  arrange(date)

## Take the sample data from the original data
readmit_data <- slice(readmit_df, 1:(n() - 300))

## Take a glimpse at the data
glimpse(readmit_data)
```

## 4.2. Training and testing splits

```{r splits}
## Set the seed
set.seed(4355)

## Create the data split
readmit_split <- initial_split(readmit_data, prop = 0.75, strata = readmitted)

## Create training and testing sets
readmit_train <- training(readmit_split)
readmit_test <- testing(readmit_split)

## Check the dimension
dim(readmit_train)
dim(readmit_test)
```

## 4.3.Load training results

Load the training results for the 8 models we trained: logistic regression, decision tree, naive bayes, k-NN, random forest, svm linear and svm RBF, and xgboost.

```{r data split}
## Setting up the control parameters
grid_ctrl <- control_grid(
  verbose = TRUE,        
  save_pred = TRUE,
  parallel_over = "everything",
  save_workflow = TRUE
)

## Load the grid results
grid_results <- read_rds(here("results", "readmit_grid_results.rds"))

```

```{r results}

grid_results
```




# 5. Modelling

Create an ensemble model with the 8 trained model.

```{r stack model}
## Create the stack object
model_stack <- 
  ## Initialize the stack
  stacks() %>% 
  ## Add candidate members
  add_candidates(grid_results) %>% 
  ## Determine how to combine their predictions
  blend_predictions(metric = metric_set(f_meas),
                    control = grid_ctrl) %>% 
  ## Fit the candidates with non-zero stacking coefficients
  fit_members()

## Save as another model object
readmit_model <- model_stack

## Create a plot of the model
autoplot(readmit_model)

```


```{r another plot}
autoplot(readmit_model, type="members")
```


# 6. Evaluate the model on the test set

Assess the performance of the ensemble model on the test set (new data).

```{r stack eval}
## Generate predictions from the test set
readmit_test_pred <- 
  readmit_test %>% 
  select(readmitted) %>% 
  bind_cols(predict(readmit_model, readmit_test, members = TRUE)) 

## Print the test predictions
readmit_test_pred


## Create a confusion matrix
readmit_test_pred %>% 
  conf_mat(estimate = .pred_class, truth = readmitted) %>% 
  autoplot()

## Let's see how the ensemble model performs
map(readmit_test_pred, f_meas_vec, truth = readmit_test_pred$readmitted) %>% 
  as_tibble()


```



# 7. Create a vetiver object

Create a Vetiver object.

```{r}
## Create a vetiver object
v_model <- vetiver::vetiver_model(
  readmit_model, 
  model_name = "readmit_pred")

## Turn on versioning explicitly for a temporary demo board
model_board <- pins::board_temp(versioned = TRUE)

## Store the model's version
vetiver::vetiver_pin_write(model_board, v_model)

## Read the vetiver model object from the board
model_board %>% vetiver::vetiver_pin_read("readmit_pred")

```


# 8. Store and version the model

Store and create different versions of the model.
```{r}
## Load the readmission model from the logistic classifier
readmit_model <- read_rds(here("results", "final_readmission_model.rds"))

## Create a vetiver object
v_model <- vetiver_model(readmit_model, "readmit_pred")

## Store model's version
model_board %>% vetiver_pin_write(v_model)

## Print both versions of the model
model_board %>% pin_versions("readmit_pred")

```

# 9. Create a REST API for deployment

Deploy the vetiver model object as a Plumber API and add a POST endpoint to make predictions.

```{r plumber API}
## Reload the stacked model object
readmit_model <- model_stack

## Create a vetiver model object
v_model <- vetiver_model(readmit_model, "readmit_pred")

## Deploy the model by creating a special Plumber router
pr() %>% 
  vetiver_api(v_model) %>% 
  pr_run(port = 8088)
```

Running the code above would launch the API in a browser.

**Steps to make predictions:**

1. After running the code chunk above, a web interface will open

2. Click on **return predictions from model using 11 features** because we want to make predictions

3. Under **POST/Predict REQUEST**, click on example. You can copy the example below to test out the API.

[
  {
    "race": "Others",
    "sex": "Female",
    "age": "<60 years",
    "hospital_stay": 7,
    "HbA1c": "Normal",
    "diabetesMed": "No",
    "admit_source": "Emerg",
    "patient_visits": 3,
    "num_medications": 20,
    "num_diagnosis": 8,
    "insulin_level": "Up"
  }
]


4. Click on the **TRY** button to get a prediction.


# 10. Create a pins board for deployment

Create a pins model board for deployment.

```{r}
## Create a pins board 
model_board <- board_folder("pins_r", versioned = TRUE)

## Write the model to a pins board
vetiver_pin_write(model_board, v_model)

## Deploy to Posit Connect
#vetiver_deploy_rsconnect(
#    board = model_board, 
#    name = "olayinka-arimoro/readmit_pred",
#    predict_args = list(debug = TRUE),
#    account = "olayinka-arimoro"
#)
```

# 11. Deploy the model via Docker
Deploy the model via Docker by building a Docker container.

```{r docker}
## Build a Docker container
vetiver_prepare_docker(model_board, "readmit_pred")

```

**Open the Docker file, copy, and paste the lines below:**
RUN mkdir -p /opt/ml/pins_r
ADD pins_r/ /opt/ml/pins_r

**Head over to the terminal tab and run**
docker build --platform linux/amd64 -t readmit .

**Once the image is built, you can launch by running**
docker run -p 8000:8000 readmit

**To stop the docker container, run:**

**Windows:** docker stop -t 60 container_id

`container_id` is the name generated by docker when you click on "container" section in docker desktop.

**Mac:** docker stop $(docker ps -a -q)

# 12. Predict from the model's endpoint

Make predictions using the vetiver model's endpoint.

```{r docker API endpoint}
## Create the endpoint
v_endpoint <- vetiver_endpoint("http://127.0.0.1:8000/predict")

## Print the model's endpoint
v_endpoint

## Create data for a new patient
new_patient <- tibble(race = "Others", sex = "Female",
                  age = "<60 years", hospital_stay = 7,
                  HbA1c = "Normal", diabetesMed = "No",
                  admit_source = "Emerg", patient_visits = 3,
                  num_medications = 20, num_diagnosis = 8,
                  insulin_level = "Up")

## Print out the new patient
new_patient

## Make the prediction
predict(v_endpoint, new_patient)

```


# 13. Compute monitoring metrics

Compute monitoring metrics.

```{r monitor metric}
## Read the model from the board
v_readmit <- vetiver_pin_read(model_board, "readmit_pred")

## Take a glimpse at the validation set
glimpse(readmit_val_set)

## You may save this validation set to the board
model_board %>% pin_write(readmit_val_set, "readmit_valset")

## Compute monitoring metrics
readmit_metrics <- 
  augment(v_readmit, new_data = readmit_val_set) %>% 
  vetiver_compute_metrics(
    date_var = date,
    period = "month",
    truth = readmitted,
    estimate = .pred_class,
    metric_set = yardstick::metric_set(f_meas, sens)
  )

## Print the metrics result
readmit_metrics

```

# 14. Pin monitoring metrics

Pin monitoring metrics to the board.

```{r monitor}
## Pin the metrics to the board
pins::pin_write(model_board, readmit_metrics, "readmit_metrics")

## Create a sample of new data
readmit_new_valset <-
  readmit_data |> 
  slice_tail(n = 190) |> 
  arrange(date)

## Compute monitoring metrics with new data
new_readmit_metrics <- 
  augment(v_readmit, new_data = readmit_new_valset) %>% 
  vetiver_compute_metrics(
    date_var = date,
    period = "month",
    truth = readmitted,
    estimate = .pred_class,
    metric_set = yardstick::metric_set(f_meas, sens)
  )

## Print the new metrics result
new_readmit_metrics

## Pin the new metrics to the board
vetiver::vetiver_pin_metrics(
  model_board, 
  new_readmit_metrics, 
  "readmit_metrics", 
  overwrite = TRUE)

```

# 15. Plot monitoring metrics

Create a plot to track the model's performance over time.

```{r monitor plot}
## Read the pinned metrics from the board
monitoring_metrics <- model_board %>% 
  pin_read("readmit_metrics")

## Plot the metrics
vetiver_plot_metrics(monitoring_metrics) +
  scale_size(range = c(2,4)) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    size = "Number of patients"
  )

```


VoilA...










